---
title: "When do SOTA Activators post their spots?"
date: now
format: gfm
fig-dpi: 600
execute: 
  warning: false
---

The goal of this project is to find the best times to be on a SOTA summit to have a summit-to-summit contact. 


```{r}
library(httr2)
library(tidyverse)
library(sf)
library(lutz)
library(hms)
library(jsonlite)

# get the most recent file
most_recent_file <- 
  list.files("data", full.names = TRUE) %>% 
  magrittr::extract(which.max(file.mtime(.)))

resp_spot <- readRDS(most_recent_file)

# getting spots is documented in the R script file
```

These data are from the 72 h leading up to `r format(file.mtime(most_recent_file), format='%Y-%m-%d %H:%M')`. The data are collected by a GitHub Action that is set to run every 72 h.

```{r}
# get associations so we can get their time zones via lat-long
req <- request("https://api2.sota.org.uk/api/")
resp_assoc <- req %>% 
  req_url_path_append("associations") %>% 
  req_perform() %>% 
  resp_body_json() 
```

First, find the time zone that each SOTA Association is in. We assume each Association has one time zone (this is probably not true) and we ignore daylight savings. 

```{r}
assoc_tbl <- 
resp_assoc %>% 
  tibble(resp = resp_assoc) %>% 
  unnest_wider(resp) 

# find the timezones for each association
assoc_tbl_tz <- 
  assoc_tbl %>% 
  mutate(activeFrom = str_replace_all(activeFrom, "T", " ")) %>% 
  mutate(date = as_datetime(activeFrom, 
                            format = "%Y-%m-%d %H:%M:%S", 
                            tz = "GMT"),
         timezone = tz_lookup_coords(lat = minLat, 
                                     lon = minLong, 
                                     method = "accurate")) 

spot_tbl <- 
resp_spot %>% 
  tibble(resp_spot = resp_spot) %>% 
  unnest_wider(resp_spot) %>% 
  select(-.)
```

Now we have the time zone for each Association, let's compute the local time that each spot was posted, and the day of the week. 

```{r}
# find the time zone for each spot, via the association that the spot is in
spot_tbl_tz <- 
spot_tbl %>% 
  left_join(assoc_tbl_tz %>% 
              select(associationCode, 
                     associationName,
                     timezone),
            join_by("associationCode")) %>% 
  mutate(timeStamp = str_replace_all(timeStamp, "T", " ")) %>% 
  mutate(timeStamp1 = as_datetime(timeStamp, 
                                format = "%Y-%m-%d %H:%M:%OS", 
                                tz = "GMT")) 

# convert UTC times from the API to local times
# we cannot have multiple time zones in one vector
# so this stays as a list
tzs <- vector("list", length = nrow(spot_tbl_tz))
for(i in 1:nrow(spot_tbl_tz)){
  tzs[[i]] <- with_tz(spot_tbl_tz$timeStamp1[i], 
                      tzone = spot_tbl_tz$timezone[i])
}

# convert time to HMS only, get weekdays
days.of.week <- weekdays(as.Date(4,"1970-01-01",tz="GMT")+0:6)
spot_tbl_tz_local <- 
  spot_tbl_tz %>% 
  mutate(local_time = map_vec(tzs,  ~as_hms(.x))) %>% 
  mutate(day_of_the_week = map_vec(tzs,  ~weekdays(.x)))  %>% 
  mutate(day_of_the_week = factor(day_of_the_week,
                                  levels = days.of.week))

# 50% of spots occur between these local times
fifty_perc <- quantile(spot_tbl_tz_local$local_time, c(0.25,  0.75))

fifty_perc_1 <- format(round_date(as.POSIXct(paste("1900-01-01 ", fifty_perc[1])), unit="1 mins"), "%H:%M")
fifty_perc_2 <- format(round_date(as.POSIXct(paste("1900-01-01 ", fifty_perc[2])), unit="1 mins"), "%H:%M")
```

Most spots are posted on Saturday. 

```{r}
#| label: fig-by-day
#| fig-cap: "Frequency of spots by day of the week"

# plot all in one histogram
spot_tbl_tz_local  %>% 
  ggplot() +
  aes(day_of_the_week) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  theme_minimal() +
  ggtitle("All spots on all modes in all Associations")
```



```{r}
#| label: fig-by-hour
#| fig-cap: "Frequency of spots by time of the day in the activator's time zone"


spot_tbl_tz_local  %>% 
  ggplot() +
  aes(local_time) +
  geom_histogram() +
  geom_vline(xintercept = fifty_perc, colour = "red") +
  theme_minimal() +
  xlab("Local time\n50% of spots occur between the red lines") +
  ggtitle("All spots on all modes in all Associations")

```

Midday is peak hour for spotting. Fifty percent of spots were posted between `r fifty_perc_1` and `r fifty_perc_2` in the activator's local time zone.

```{r}
#| label: fig-by-hour-day
#| fig-cap: "Frequency of spots by time in the activator's time zone and day of the week"

small_x_labels <- 3

# plot all in one histogram
spot_tbl_tz_local  %>% 
ggplot() +
  aes(local_time) +
  geom_histogram() +
  geom_vline(xintercept = fifty_perc, colour = "red") +
  theme_minimal() +
  facet_wrap(~ day_of_the_week, scales = "free_y") +
  xlab("Local time\n50% of spots occur between the red lines") +
  ggtitle("All spots on all modes in all Associations") +
  theme(axis.text=element_text(size = small_x_labels))
```

Midday seems to be preferred regardless of the day of the week.

```{r}
#| label: fig-by-hour-by-association
#| fig-cap: "Frequency of spots by time in the activator's time zone for each Association"

spot_tbl_tz_local %>% 
  add_count(associationName) %>% 
  filter(n >= 10) %>% 
  ggplot() +
  aes(local_time) +
  geom_histogram() +
  geom_vline(xintercept = fifty_perc, colour = "red") +
  theme_minimal() +
  xlab("Local time\n50% of spots occur between the red lines") +
  facet_wrap(~ associationName, scales = "free_y") +
  ggtitle("Spots by Association (for Assocs with more than 10 spots)") +
  theme(axis.text=element_text(size = small_x_labels))
```

Activators in Japan are spotting earlier than most others. Activators in Spain are generally later than average. Within the US, Arizona is where the early spotting is, and Virginia do the late shift. 

```{r}
#| label: fig-time-and-mode
#| fig-cap: "Frequency of spots by time in the activator's time zone and mode"

spot_tbl_tz_local  %>% 
  mutate(mode = toupper(mode)) %>% 
  add_count(mode) %>% 
  filter(n >= 10) %>% 
  ggplot() +
  aes(local_time) +
  geom_histogram() +
  geom_vline(xintercept = fifty_perc, colour = "red") +
  theme_minimal() +
  xlab("Local time\n50% of spots occur between the red lines") +
  facet_wrap(~ mode, scales = "free_y") +
  ggtitle("Spots by Mode (for Modes with more than 10 spots)") +
  theme(axis.text=element_text(size = small_x_labels))
```

Activators on FM tend to be earlier than those on SSB. Perhaps many of these are the same people, starting their activation on FM, then switching to SSB.


```{r}
#| label: fig-time-and-summit
#| fig-cap: "Frequency of spots by time in the activator's time zone and summit"


spot_tbl_tz_local %>% 
  add_count(summitCode) %>% 
  filter(n >= 10) %>% 
  ggplot() +
  aes(local_time) +
  geom_histogram() +
  geom_vline(xintercept = fifty_perc, colour = "red") +
  theme_minimal() +
  xlab("Local time\n50% of spots occur between the red lines") +
  facet_wrap(~ summitCode, scales = "free_y") +
  theme(axis.text=element_text(size = small_x_labels))
```


```{r}
#| label: fig-time-and-activator
#| fig-cap: "Frequency of spots by time in the activator's time zone for activators with more than ten spots"


spot_tbl_tz_local %>% 
  add_count(activatorCallsign) %>% 
  filter(n >= 10) %>% 
  ggplot() +
  aes(local_time) +
  geom_histogram() +
  geom_vline(xintercept = fifty_perc, colour = "red") +
  theme_minimal() +
  xlab("Local time\n50% of spots occur between the red lines") +
  facet_wrap(~ activatorCallsign, scales = "free_y") +
  ggtitle("Spots by Activator (for Activators with more than 10 spots)") +
  theme(axis.text=element_text(size = small_x_labels))

```






